---
title: "Explanatory Analysis"
author: "Ellie Kuskie"
format: pdf
editor: visual
---

## Getting in The Data and Wrangling

```{r}
# load in necessary packages
library("tidyverse")
library("brms")
library("ggeffects")
library("readxl")
library("RColorBrewer")
```

```{r}
# read in the data from the excel file
Poll_Data <- read_xlsx("Ellie_ChiSquare.xlsx", sheet = 1)
# make sure the data file is in correctly
head(Poll_Data)
```

```{r}
# Create new dataframe for just flower shape 
Fl_Shape <- Poll_Data %>% 
  # used the select function to keep only 4 columns
  select(PollinatorOrder, PollinatorFamily, PollGenusSpeciesUSE, FlowerShape) %>% 
  # Get rid of flowers with unknown shape
  filter(FlowerShape != "U")

# remove uncessary columns to just focus on pollinator order
Fl_Shape_Order <- Fl_Shape %>% 
  select(PollinatorOrder, FlowerShape) 

# data frame of 2 flower shapes I want to focus on (removed other flower shapes)
Fl_Shape_Final <- Fl_Shape_Order %>%  
  group_by(PollinatorOrder, FlowerShape) %>% 
  summarize(n = n()) %>% 
  mutate(total = sum(n),
         prop = n / sum(n)) %>% 
  ungroup() %>% 
  filter(FlowerShape == "Dish" | FlowerShape == "Tube")

# Create new dataframe for just tubular flowers
Tube_Fl_Order <- Fl_Shape_Final %>% 
  filter(FlowerShape == "Tube")

# Create new dataframe for just dish flowers
Dish_Fl_Order <- Fl_Shape_Final %>% 
  filter(FlowerShape == "Dish")

Dish_Fl_Order
```

## Statistical Analysis

```{r}
# Create new dataframe for just tubular flowers
Tube_Fl_Order <- Fl_Shape_Final %>% 
  filter(FlowerShape == "Tube")
# Print dataframe to see if it workde correctly
Tube_Fl_Order

m.Tube_Fl_Order <- brm(data = Tube_Fl_Order,
   # Made a beta_binomial distribution and used link = "identity" to not have to back transform values
   family = beta_binomial(link = "identity", link_phi = "identity"),
   # categorical model using pollinator order
   n | trials(total) ~ 0 + PollinatorOrder,
   # iterations and warmup alongside number of chains
   iter = 4000, warmup = 2000, chains = 4, cores = 4,
   # setting priors reletively broadly
   prior = c(prior(beta(2, 6), class = "b", dpar = "mu", lb = 0, ub = 1),
            prior(exponential(0.001), class = "phi", lb = 0)),
   file = "output/m.Tube_Fl_Order"
)

# viewing results of model
summary(m.Tube_Fl_Order)
# show posterior plots and chains to assess model 
plot(m.Tube_Fl_Order)
```

```{r}
# Create new dataframe for just dish flowers
Dish_Fl_Order <- Fl_Shape_Final %>% 
  filter(FlowerShape == "Dish")
# Print dataframe to see if it workde correctly
Dish_Fl_Order

m.Dish_Fl_Order <- brm(data = Dish_Fl_Order,
   # Made a beta_binomial distribution and used link = "identity" to not have to back transform values
   family = beta_binomial(link = "identity", link_phi = "identity"),
   # categorical model using pollinator order
   n | trials(total) ~ 0 + PollinatorOrder,
   # iterations and warmup alongside number of chains
   iter = 4000, warmup = 2000, chains = 4, cores = 4,
   # setting priors reletively broadly
   prior = c(prior(beta(2, 6), class = "b", dpar = "mu", lb = 0, ub = 1),
            prior(exponential(0.001), class = "phi", lb = 0)),
   file = "output/m.Dish_Fl_Order"
)

# viewing results of model
summary(m.Dish_Fl_Order)
# show posterior plots and chains to assess model 
plot(m.Dish_Fl_Order)
```

For both of these models we can see that the Rhat values are equal to 1, the posteriors all have one clear hump, and the chains converge. With this we can conclude that the model ran well.

Biologically this means that 16% of Fly visits will be to tubular flowers with a 95% credible interval from \[11% - 23%\], 37% of Bee visits will be to tubular flowers with a 95% credible interval from \[30% - 43%\], and 41% of Butterfly/Moth visits will be to tubular flowers with a 95% credible interval from \[28% - 54%\].

Biologically this means that 66% of Fly visits will be to dish flowers with a 95% credible interval from \[45% - 75%\], 49% of Bee visits will be to dish flowers with a 95% credible interval from \[33% - 46%\], and 40% of Butterfly/Moth visits will be to tubular flowers with a 95% credible interval from \[25% - 55%\].

## Graphs

```{r}
# create graph for visitation frequency on tubular flowers
TubeGraph <- ggplot(Tube_Fl_Order, aes(PollinatorOrder, y = prop)) +
  # make it a bar plot and set fill = PollinatorOrder for color later
  geom_bar(stat = "identity", mapping = aes(fill = PollinatorOrder)) +
  # Change x and y axis 
  labs(x = "Pollinator Order",
       y = "Visitation Frequency") +
  # get rid of gap at y = 0 and make y-axis percentage
  scale_y_continuous(expand = c(0, 0), labels = scales::percent, limits = c(NA, 1)) +
  # change theme
  theme_bw() +
  # remove legend
  theme(legend.position = "none") +
  # change colors of bars to look nicer
  scale_fill_brewer(palette = "Accent")
  
# print out graph
TubeGraph
```

```{r}
# create graph for visitation frequency on dish flowers
DishGraph <- ggplot(Dish_Fl_Order, aes(PollinatorOrder, y = prop)) +
  # make it a bar plot and set fill = PollinatorOrder for color later
  geom_bar(stat = "identity", mapping = aes(fill = PollinatorOrder)) +
  # Change x and y axis 
  labs(x = "Pollinator Order",
       y = "Visitation Frequency") +
  # get rid of gap at y = 0 and make y-axis percentage
  scale_y_continuous(expand = c(0, 0), labels = scales::percent, limits = c(NA, 1)) +
  # change theme
  theme_bw() +
  # remove legend
  theme(legend.position = "none") +
  # change colors of bars to look nicer
  scale_fill_brewer(palette = "Accent")
  
# print out graph
DishGraph
```

Based on these 2 figures you can see how Diptera or Flies have a strong tendency to visit flowers that are dish shaped. They also have very little tendency to visit tubular flowers when compared to the other orders examined. With Butterflies/moths we can see that they have the exact same visitation frequency to dish and tube flowers; this could indicate a lack of preference, however out data for this order is a bit limited so perhaps with more data a pattern could emerge. With Hymenoptera or Bees we can see that they tend to visit dish flowers a bit more frequently than tubular flowers; this could be due to their mouthparts being more adapted for flowers with less of a distance to nectar (similar to flies). Perhaps it is different for different families of bees, and further analysis could help us to understand this.
